Federated Learning's Blessing:\\ FedAvg has Linear Speedup

### 2020-06-10 source/app_sgd_scvx_smooth.tex
\ov{v}, \ov{w}, \vw

	1. Quadratic upper bound: $0\leq F(w)-F(w')-\langle\nabla F(w'),w-w'\rangle\leq\frac{L}{2}\|w-w'\|^{2}$.

	2. Coercivity: $\frac{1}{L}\|\nabla F(w)-\nabla F(w')\|^{2}\leq\langle\nabla F(w)-\nabla F(w'),w-w'\rangle$.

	3. Lower bound: $F(w)\geq F(w')+\langle\nabla F(w'),w-w'\rangle+\frac{1}{2L}\|\nabla F(w)-\nabla F(w')\|^{2}$.
	In particular, $\|\nabla F(w)\|^{2}\leq2L(F(w)-F(w^{\ast}))$.

	4. Optimality gap: $F(w)-F(w^{\ast})\leq$$\langle\nabla F(w),w-w^{\ast}\rangle$.
\end{lemma}
%
\begin{lemma}
	Let $F$ be a $\mu$-strongly convex function. Then 
	\begin{align*}
	F(w) & \leq F(w')+\langle\nabla F(w'),w-w'\rangle+\frac{1}{2\mu}\|\nabla F(w)-\nabla F(w')\|^{2}\\
	F(w)-F(w^{\ast}) & \leq\frac{1}{2\mu}\|\nabla F(w)\|^{2}
	\end{align*}
\end{lemma}

@@

\begin{proof}
	The first step is to observe that the $L$-smoothness of $F$ provides
	the upper bound
	\begin{align*}
	\mathbb{E}(F(\overline{w}_{t}))-F^{\ast} & =\mathbb{E}(F(\overline{w}_{t})-F(w^{\ast}))\\
	& \leq\frac{L}{2}\mathbb{E}\|\overline{w}_{t}-w^{\ast}\|^{2}
	\end{align*}
	and bound $\mathbb{E}\|\overline{w}_{t}-w^{\ast}\|^{2}$.

	Our main step is to prove the bound 
	\begin{align*}
	\mathbb{E}\|\overline{w}_{t+1}-w^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})\mathbb{E}\|\overline{w}_{t}-w^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+5E^{2}L\alpha_{t}^{3}G^{2}
	\end{align*}

	We have 
	\begin{align*}
	\|\overline{w}_{t+1}-w^{\ast}\|^{2} & =\|(\overline{w}_{t}-\alpha_{t}g_{t})-w^{\ast}\|^{2}\\
	& =\|(\overline{w}_{t}-\alpha_{t}\overline{g}_{t}-w^{\ast})-\alpha_{t}(g_{t}-\overline{g}_{t})\|^{2}\\
	& =A_{1}+A_{2}+A_{3}
	\end{align*}
	where 
	\begin{align*}
	A_{1} & =\|\overline{w}_{t}-w^{\ast}-\alpha_{t}\overline{g}_{t}\|^{2}\\
	A_{2} & =2\alpha_{t}\langle\overline{w}_{t}-w^{\ast}-\alpha_{t}\overline{g}_{t},\overline{g}_{t}-g_{t}\rangle\\
	A_{3} & =\alpha_{t}^{2}\|g_{t}-\overline{g}_{t}\|^{2}
	\end{align*}
	$\mathbb{E}A_{2}=0$ by definition of $g_{t}$ and $\overline{g}_{t}$,
	while for $A_{3}$ we have
	\begin{align*}
	\alpha_{t}^{2}\mathbb{E}\|g_{t}-\overline{g}_{t}\|^{2} & =\alpha_{t}^{2}\mathbb{E}\|g_{t}-\mathbb{E}g_{t}\|^{2}=\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\|g_{t,k}-\mathbb{E}g_{t,k}\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
	\end{align*}
	again by Jensen's inequality and using the independence of $g_{t,k},g_{t,k'}$.

	Next we bound $A_{1}$: 
	\begin{align*}
	\|\overline{w}_{t}-w^{\ast}-\alpha_{t}\overline{g}_{t}\|^{2} & =\|\overline{w}_{t}-w^{\ast}\|^{2}+2\langle\overline{w}_{t}-w^{\ast},-\alpha_{t}\overline{g}_{t}\rangle+\|\alpha_{t}\overline{g}_{t}\|^{2}
	\end{align*}
	and we will show that the third term $\|\alpha_{t}\overline{g}_{t}\|^{2}$
	can be canceled by an upper bound of the second term. %
	\begin{comment}
	The last term is straightforward to bound by the convexity of $\|\cdot\|^{2}$
	and $L$-smoothness of $F_{k}$,
	\begin{align*}
	\alpha_{t}^{2}\|\overline{g}_{t}\|^{2} & \leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\|\nabla F_{k}(w_{t}^{k})\|^{2}\leq2L\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}(F_{k}(w_{t}^{k})-F_{k}^{\ast})
	\end{align*}
	or 
	\begin{align*}
	\alpha_{t}^{2}\|\overline{g}_{t}\|^{2} & \leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\|\nabla F_{k}(w_{t}^{k})\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\mathbb{E}\|\nabla F_{k}(w_{t}^{k},\xi_{t}^{k})\|^{2}\leq\alpha_{t}^{2}G^{2}
	\end{align*}
	\end{comment}

	Now 
	\begin{align*}
	& -2\alpha_{t}\langle\overline{w}_{t}-w^{\ast},\overline{g}_{t}\rangle\\
	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\overline{w}_{t}-w^{\ast},\nabla F_{k}(w_{t}^{k})\rangle\\
	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\overline{w}_{t}-w_{t}^{k},\nabla F_{k}(w_{t}^{k})\rangle-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle w_{t}^{k}-w^{\ast},\nabla F_{k}(w_{t}^{k})\rangle\\
	& \leq-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\overline{w}_{t}-w_{t}^{k},\nabla F_{k}(w_{t}^{k})\rangle+2\alpha_{t}\sum_{k=1}^{N}p_{k}(F_{k}(w^{\ast})-F_{k}(w_{t}^{k}))-\alpha_{t}\mu\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-w^{\ast}\|^{2}\\
	& \leq2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(w_{t}^{k})-F_{k}(\overline{w}_{t})+\frac{L}{2}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+F_{k}(w^{\ast})-F_{k}(w_{t}^{k})\right]-\alpha_{t}\mu\|\sum_{k=1}^{N}p_{k}w_{t}^{k}-w^{\ast}\|^{2}\\
	& =\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(w^{\ast})-F_{k}(\overline{w}_{t})\right]-\alpha_{t}\mu\|\overline{w}_{t}-w^{\ast}\|^{2}
	\end{align*}
	For the second term, which is negative, we can ignore it, but this
	yields a suboptimal bound that fails to provide the desired linear
	speedup. Instead, we upper bound it using the following derivation:
	\begin{align*}
	& 2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(w^{\ast})-F_{k}(\overline{w}_{t})\right]\\
	& \leq2\alpha_{t}\left[F(\overline{w}_{t+1})-F(\overline{w}_{t})\right]\\
	& \leq2\alpha_{t}\mathbb{E}\langle\nabla F(\overline{w}_{t}),\overline{w}_{t+1}-\overline{w}_{t}\rangle+\alpha_{t}L\mathbb{E}\|\overline{w}_{t+1}-\overline{w}_{t}\|^{2}\\
	& =-2\alpha_{t}^{2}\mathbb{E}\langle\nabla F(\overline{w}_{t}),g_{t}\rangle+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
	& =-2\alpha_{t}^{2}\mathbb{E}\langle\nabla F(\overline{w}_{t}),\overline{g}_{t}\rangle+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
	& =-\alpha_{t}^{2}\left[\|\nabla F(\overline{w}_{t})\|^{2}+\|\overline{g}_{t}\|^{2}-\|\nabla F(\overline{w}_{t})-\overline{g}_{t}\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
	& =-\alpha_{t}^{2}\left[\|\nabla F(\overline{w}_{t})\|^{2}+\|\overline{g}_{t}\|^{2}-\|\nabla F(\overline{w}_{t})-\sum_{k}p_{k}\nabla F(w_{t}^{k})\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
	& \leq-\alpha_{t}^{2}\left[\|\nabla F(\overline{w}_{t})\|^{2}+\|\overline{g}_{t}\|^{2}-\sum_{k}p_{k}\|\nabla F(\overline{w}_{t})-\nabla F(w_{t}^{k})\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
	& \leq-\alpha_{t}^{2}\left[\|\nabla F(\overline{w}_{t})\|^{2}+\|\overline{g}_{t}\|^{2}-L^{2}\sum_{k}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
	& \leq-\alpha_{t}^{2}\|\overline{g}_{t}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}-\alpha_{t}^{2}\|\nabla F(\overline{w}_{t})\|^{2}
	\end{align*}
	where we have used the smoothness of $F$ twice.

	Note that the term $-\alpha_{t}^{2}\|\overline{g}_{t}\|^{2}$ exactly
	cancels the $\alpha_{t}^{2}\|\overline{g}_{t}\|^{2}$ in the bound
	for $A_{1}$, so that plugging in the bound for $-2\alpha_{t}\langle\overline{w}_{t}-w^{\ast},\overline{g}_{t}\rangle$,
	we have so far proved 
	\begin{align*}
	\mathbb{E}\|\overline{w}_{t+1}-w^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})\|\overline{w}_{t}-w^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
	& +\alpha_{t}^{2}L^{2}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}-\alpha_{t}^{2}\|\nabla F(\overline{w}_{t})\|^{2}
	\end{align*}
	The term $\mathbb{E}\|g_{t}\|^{2}\leq G^{2}$ by assumption.

	Now we bound $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}$.
	Since communication is done every $E$ steps, for any $t\geq0$, we
	can find a $t_{0}\leq t$ such that $t-t_{0}\leq E-1$ and $w_{t_{0}}^{k}=\overline{w}_{t_{0}}$for
	all $k$. Moreover, using $\alpha_{t}$ is non-increasing and $\alpha_{t_{0}}\leq2\alpha{}_{t}$
	for any $t-t_{0}\leq E-1$, we have 
	\begin{align*}
	& \mathbb{E}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}\\
	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-\overline{w}_{t_{0}}-(\overline{w}_{t}-\overline{w}_{t_{0}})\|^{2}\\
	& \leq\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-\overline{w}_{t_{0}}\|^{2}\\
	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-w_{t_{0}}^{k}\|^{2}\\
	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|-\sum_{i=t_{0}}^{t-1}\alpha_{i}g_{i,k}\|^{2}\\
	& \leq2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}E\alpha_{i}^{2}\|g_{i,k}\|^{2}\\
	& \leq2\sum_{k=1}^{N}p_{k}E^{2}\alpha_{t_{0}}^{2}G^{2}\\
	& \leq4E^{2}\alpha_{t}^{2}G^{2}
	\end{align*}
	\begin{comment}
	Alternatively, we can also write
	\begin{align*}
	\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2} & =\sum_{k=1}^{N}p_{k}\|\overline{w}_{t-1}-\alpha_{t-1}g_{t-1}-w_{t-1}^{k}+\alpha_{t-1}g_{t-1,k}\|^{2}\\
	& \leq2\sum_{k=1}^{N}p_{k}\left(\|\overline{w}_{t-1}-w_{t-1}^{k}\|^{2}+\|\alpha_{t-1}g_{t-1}-\alpha_{t-1}g_{t-1,k}\|^{2}\right)
	\end{align*}
	and write 
	\begin{align*}
	\sum_{k}p_{k}\|g_{t-1,k}-g_{t-1}\|^{2}\leq\sum_{k}p_{k}\|g_{t-1,k}\|^{2} & =\sum_{k}p_{k}\|g_{t-1,k}-\nabla F_{k}(w_{t-1}^{k})+\nabla F_{k}(w_{t-1}^{k})\|^{2}\\
	& \leq2\sum_{k}p_{k}\|g_{t-1,k}-\nabla F_{k}(w_{t-1}^{k})\|^{2}+2\sum_{k}p_{k}\|\nabla F_{k}(w_{t-1}^{k})\|^{2}\\
	& \leq2\sum_{k}p_{k}\sigma_{k}^{2}+2\sum_{k}p_{k}\|\nabla F_{k}(w_{t-1}^{k})-\nabla F_{k}(\overline{w}_{t-1})+\nabla F_{k}(\overline{w}_{t-1})\|^{2}\\
	& \leq2\sum_{k}p_{k}\sigma_{k}^{2}+4\sum_{k}p_{k}\left(\|\nabla F_{k}(w_{t-1}^{k})-\nabla F_{k}(\overline{w}_{t-1})\|^{2}+\|\nabla F_{k}(\overline{w}_{t-1})\|^{2}\right)\\
	& \leq2\sigma^{2}+4\sum_{k}p_{k}\left(L^{2}\|\overline{w}_{t-1}-w_{t-1}^{k}\|^{2}+\|\nabla F_{k}(\overline{w}_{t-1})\|^{2}\right)
	\end{align*}
	so that 
	\begin{align*}
	\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2} & \leq2(1+4L^{2}\alpha_{t-1}^{2})\sum_{k=1}^{N}p_{k}\|\overline{w}_{t-1}-w_{t-1}^{k}\|^{2}+4\sigma^{2}\alpha_{t-1}^{2}+8\alpha_{t-1}^{2}\sum_{k}p_{k}\|\nabla F_{k}(\overline{w}_{t-1})\|^{2}
	\end{align*}
	and taking expectation, we have 
	\begin{align*}
	\mathbb{E}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2} & \leq2(1+4L^{2}\alpha_{t-1}^{2})\mathbb{E}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t-1}-w_{t-1}^{k}\|^{2}+4\alpha_{t-1}^{2}(\sigma^{2}+2G^{2})
	\end{align*}
	Using this recusive relation at most $E$ times to arrive at a communication
	round and that $\alpha_{t}L\leq1/8$ for all $t$, we have 
	\begin{align*}
	\mathbb{E}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}\leq E\frac{1}{L^{2}}(\sigma^{2}+2G^{2})
	\end{align*}
	which gives 
	\end{comment}
	{}

>>>>

	1. Quadratic upper bound: $0\leq F(\vw)-F(\vw')-\langle\nabla F(\vw'),\vw-\vw'\rangle\leq\frac{L}{2}\|\vw-\vw'\|^{2}$.

	2. Coercivity: $\frac{1}{L}\|\nabla F(\vw)-\nabla F(\vw')\|^{2}\leq\langle\nabla F(\vw)-\nabla F(\vw'),\vw-\vw'\rangle$.

	3. Lower bound: $F(\vw)\geq F(\vw')+\langle\nabla F(\vw'),\vw-\vw'\rangle+\frac{1}{2L}\|\nabla F(\vw)-\nabla F(\vw')\|^{2}$.
	In particular, $\|\nabla F(\vw)\|^{2}\leq2L(F(\vw)-F(\vw^{\ast}))$.

	4. Optimality gap: $F(\vw)-F(\vw^{\ast})\leq$$\langle\nabla F(\vw),\vw-\vw^{\ast}\rangle$.
\end{lemma}
%
\begin{lemma}
	Let $F$ be a $\mu$-strongly convex function. Then 
	\begin{align*}
	F(\vw) & \leq F(\vw')+\langle\nabla F(\vw'),\vw-\vw'\rangle+\frac{1}{2\mu}\|\nabla F(\vw)-\nabla F(\vw')\|^{2}\\
	F(\vw)-F(\vw^{\ast}) & \leq\frac{1}{2\mu}\|\nabla F(\vw)\|^{2}
	\end{align*}
\end{lemma}

@@

\begin{proof}
	The first step is to observe that the $L$-smoothness of $F$ provides
	the upper bound
	\begin{align*}
	\mathbb{E}(F(\ov{w}_{t}))-F^{\ast} & =\mathbb{E}(F(\ov{w}_{t})-F(\vw^{\ast}))\\
	& \leq\frac{L}{2}\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}
	\end{align*}
	and bound $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}$.

	Our main step is to prove the bound 
	\begin{align*}
	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+5E^{2}L\alpha_{t}^{3}G^{2}
	\end{align*}

	We have 
	\begin{align*}
	\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & =\|(\ov{w}_{t}-\alpha_{t}g_{t})-\vw^{\ast}\|^{2}\\
	& =\|(\ov{w}_{t}-\alpha_{t}\overline{g}_{t}-\vw^{\ast})-\alpha_{t}(g_{t}-\overline{g}_{t})\|^{2}\\
	& =A_{1}+A_{2}+A_{3}
	\end{align*}
	where 
	\begin{align*}
	A_{1} & =\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\overline{g}_{t}\|^{2}\\
	A_{2} & =2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\overline{g}_{t},\overline{g}_{t}-g_{t}\rangle\\
	A_{3} & =\alpha_{t}^{2}\|g_{t}-\overline{g}_{t}\|^{2}
	\end{align*}
	$\mathbb{E}A_{2}=0$ by definition of $g_{t}$ and $\overline{g}_{t}$,
	while for $A_{3}$ we have
	\begin{align*}
	\alpha_{t}^{2}\mathbb{E}\|g_{t}-\overline{g}_{t}\|^{2} & =\alpha_{t}^{2}\mathbb{E}\|g_{t}-\mathbb{E}g_{t}\|^{2}=\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\|g_{t,k}-\mathbb{E}g_{t,k}\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
	\end{align*}
	again by Jensen's inequality and using the independence of $g_{t,k},g_{t,k'}$.

	Next we bound $A_{1}$: 
	\begin{align*}
	\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\overline{g}_{t}\|^{2} & =\|\ov{w}_{t}-\vw^{\ast}\|^{2}+2\langle\ov{w}_{t}-\vw^{\ast},-\alpha_{t}\overline{g}_{t}\rangle+\|\alpha_{t}\overline{g}_{t}\|^{2}
	\end{align*}
	and we will show that the third term $\|\alpha_{t}\overline{g}_{t}\|^{2}$
	can be canceled by an upper bound of the second term. %
	\begin{comment}
	The last term is straightforward to bound by the convexity of $\|\cdot\|^{2}$
	and $L$-smoothness of $F_{k}$,
	\begin{align*}
	\alpha_{t}^{2}\|\overline{g}_{t}\|^{2} & \leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\|\nabla F_{k}(\vw_{t}^{k})\|^{2}\leq2L\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}(F_{k}(\vw_{t}^{k})-F_{k}^{\ast})
	\end{align*}
	or 
	\begin{align*}
	\alpha_{t}^{2}\|\overline{g}_{t}\|^{2} & \leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\|\nabla F_{k}(\vw_{t}^{k})\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\mathbb{E}\|\nabla F_{k}(\vw_{t}^{k},\xi_{t}^{k})\|^{2}\leq\alpha_{t}^{2}G^{2}
	\end{align*}
	\end{comment}

	Now 
	\begin{align*}
	& -2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast},\overline{g}_{t}\rangle\\
	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw^{\ast},\nabla F_{k}(\vw_{t}^{k})\rangle\\
	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw_{t}^{k},\nabla F_{k}(\vw_{t}^{k})\rangle-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle w_{t}^{k}-\vw^{\ast},\nabla F_{k}(\vw_{t}^{k})\rangle\\
	& \leq-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw_{t}^{k},\nabla F_{k}(\vw_{t}^{k})\rangle+2\alpha_{t}\sum_{k=1}^{N}p_{k}(F_{k}(\vw^{\ast})-F_{k}(\vw_{t}^{k}))-\alpha_{t}\mu\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-\vw^{\ast}\|^{2}\\
	& \leq2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw_{t}^{k})-F_{k}(\ov{w}_{t})+\frac{L}{2}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+F_{k}(\vw^{\ast})-F_{k}(\vw_{t}^{k})\right]-\alpha_{t}\mu\|\sum_{k=1}^{N}p_{k}w_{t}^{k}-\vw^{\ast}\|^{2}\\
	& =\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]-\alpha_{t}\mu\|\ov{w}_{t}-\vw^{\ast}\|^{2}
	\end{align*}
	For the second term, which is negative, we can ignore it, but this
	yields a suboptimal bound that fails to provide the desired linear
	speedup. Instead, we upper bound it using the following derivation:
	\begin{align*}
	& 2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]\\
	& \leq2\alpha_{t}\left[F(\ov{w}_{t+1})-F(\ov{w}_{t})\right]\\
	& \leq2\alpha_{t}\mathbb{E}\langle\nabla F(\ov{w}_{t}),\ov{w}_{t+1}-\ov{w}_{t}\rangle+\alpha_{t}L\mathbb{E}\|\ov{w}_{t+1}-\ov{w}_{t}\|^{2}\\
	& =-2\alpha_{t}^{2}\mathbb{E}\langle\nabla F(\ov{w}_{t}),g_{t}\rangle+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
	& =-2\alpha_{t}^{2}\mathbb{E}\langle\nabla F(\ov{w}_{t}),\overline{g}_{t}\rangle+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
	& =-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\overline{g}_{t}\|^{2}-\|\nabla F(\ov{w}_{t})-\overline{g}_{t}\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
	& =-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\overline{g}_{t}\|^{2}-\|\nabla F(\ov{w}_{t})-\sum_{k}p_{k}\nabla F(\vw_{t}^{k})\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
	& \leq-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\overline{g}_{t}\|^{2}-\sum_{k}p_{k}\|\nabla F(\ov{w}_{t})-\nabla F(\vw_{t}^{k})\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
	& \leq-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\overline{g}_{t}\|^{2}-L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
	& \leq-\alpha_{t}^{2}\|\overline{g}_{t}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}-\alpha_{t}^{2}\|\nabla F(\ov{w}_{t})\|^{2}
	\end{align*}
	where we have used the smoothness of $F$ twice.

	Note that the term $-\alpha_{t}^{2}\|\overline{g}_{t}\|^{2}$ exactly
	cancels the $\alpha_{t}^{2}\|\overline{g}_{t}\|^{2}$ in the bound
	for $A_{1}$, so that plugging in the bound for $-2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast},\overline{g}_{t}\rangle$,
	we have so far proved 
	\begin{align*}
	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
	& +\alpha_{t}^{2}L^{2}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}-\alpha_{t}^{2}\|\nabla F(\ov{w}_{t})\|^{2}
	\end{align*}
	The term $\mathbb{E}\|g_{t}\|^{2}\leq G^{2}$ by assumption.

	Now we bound $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}$.
	Since communication is done every $E$ steps, for any $t\geq0$, we
	can find a $t_{0}\leq t$ such that $t-t_{0}\leq E-1$ and $w_{t_{0}}^{k}=\ov{w}_{t_{0}}$for
	all $k$. Moreover, using $\alpha_{t}$ is non-increasing and $\alpha_{t_{0}}\leq2\alpha{}_{t}$
	for any $t-t_{0}\leq E-1$, we have 
	\begin{align*}
	& \mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\\
	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-\ov{w}_{t_{0}}-(\ov{w}_{t}-\ov{w}_{t_{0}})\|^{2}\\
	& \leq\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-\ov{w}_{t_{0}}\|^{2}\\
	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-\vw_{t_{0}}^{k}\|^{2}\\
	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|-\sum_{i=t_{0}}^{t-1}\alpha_{i}g_{i,k}\|^{2}\\
	& \leq2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}E\alpha_{i}^{2}\|g_{i,k}\|^{2}\\
	& \leq2\sum_{k=1}^{N}p_{k}E^{2}\alpha_{t_{0}}^{2}G^{2}\\
	& \leq4E^{2}\alpha_{t}^{2}G^{2}
	\end{align*}
	\begin{comment}
	Alternatively, we can also write
	\begin{align*}
	\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} & =\sum_{k=1}^{N}p_{k}\|\ov{w}_{t-1}-\alpha_{t-1}g_{t-1}-\vw_{t-1}^{k}+\alpha_{t-1}g_{t-1,k}\|^{2}\\
	& \leq2\sum_{k=1}^{N}p_{k}\left(\|\ov{w}_{t-1}-\vw_{t-1}^{k}\|^{2}+\|\alpha_{t-1}g_{t-1}-\alpha_{t-1}g_{t-1,k}\|^{2}\right)
	\end{align*}
	and write 
	\begin{align*}
	\sum_{k}p_{k}\|g_{t-1,k}-g_{t-1}\|^{2}\leq\sum_{k}p_{k}\|g_{t-1,k}\|^{2} & =\sum_{k}p_{k}\|g_{t-1,k}-\nabla F_{k}(\vw_{t-1}^{k})+\nabla F_{k}(\vw_{t-1}^{k})\|^{2}\\
	& \leq2\sum_{k}p_{k}\|g_{t-1,k}-\nabla F_{k}(\vw_{t-1}^{k})\|^{2}+2\sum_{k}p_{k}\|\nabla F_{k}(\vw_{t-1}^{k})\|^{2}\\
	& \leq2\sum_{k}p_{k}\sigma_{k}^{2}+2\sum_{k}p_{k}\|\nabla F_{k}(\vw_{t-1}^{k})-\nabla F_{k}(\ov{w}_{t-1})+\nabla F_{k}(\ov{w}_{t-1})\|^{2}\\
	& \leq2\sum_{k}p_{k}\sigma_{k}^{2}+4\sum_{k}p_{k}\left(\|\nabla F_{k}(\vw_{t-1}^{k})-\nabla F_{k}(\ov{w}_{t-1})\|^{2}+\|\nabla F_{k}(\ov{w}_{t-1})\|^{2}\right)\\
	& \leq2\sigma^{2}+4\sum_{k}p_{k}\left(L^{2}\|\ov{w}_{t-1}-\vw_{t-1}^{k}\|^{2}+\|\nabla F_{k}(\ov{w}_{t-1})\|^{2}\right)
	\end{align*}
	so that 
	\begin{align*}
	\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} & \leq2(1+4L^{2}\alpha_{t-1}^{2})\sum_{k=1}^{N}p_{k}\|\ov{w}_{t-1}-\vw_{t-1}^{k}\|^{2}+4\sigma^{2}\alpha_{t-1}^{2}+8\alpha_{t-1}^{2}\sum_{k}p_{k}\|\nabla F_{k}(\ov{w}_{t-1})\|^{2}
	\end{align*}
	and taking expectation, we have 
	\begin{align*}
	\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} & \leq2(1+4L^{2}\alpha_{t-1}^{2})\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t-1}-\vw_{t-1}^{k}\|^{2}+4\alpha_{t-1}^{2}(\sigma^{2}+2G^{2})
	\end{align*}
	Using this recusive relation at most $E$ times to arrive at a communication
	round and that $\alpha_{t}L\leq1/8$ for all $t$, we have 
	\begin{align*}
	\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\leq E\frac{1}{L^{2}}(\sigma^{2}+2G^{2})
	\end{align*}
	which gives 
	\end{comment}
	{}
